---
output: html_document
---

# Descriptive Statistics

## Summary function

```{r}
library(dplyr)
data("mtcars")
summary(select(mtcars, mpg, cyl, disp))
```

## Hmisc pacakge

```{r}
#install.packages('Hmisc')
Hmisc::describe(select(mtcars, mpg, cyl, disp))
```

## pastecs package

```{r}
# install.packages('pastecs')
pastecs::stat.desc(select(mtcars, mpg, cyl, disp))
```

## psych package

```{r}
#install.packages('psych')
psych::describe(select(mtcars, mpg, cyl, disp))
```

## Descriptive statistics by group

### Grouping by 1 var

```{r}
by(mtcars$mpg,
   list(Has_gear = mtcars$am),
   summary)
```

### Grouping by 2 var

```{r}
by(select(mtcars, mpg, cyl), #data
   list(Has_gear = mtcars$am, #grouping
        Engine = mtcars$vs),
   summary) #function
```

# Hypothesis testing

## One sample test

```{r}

library(dplyr) 
 
# Set seed for reproducibility 
set.seed(12345) 
 
# Create a dummy dataset 
n <- 1000 
group <- rep(0:1, length.out = n) 
score <- 50 + group * 10 + rnorm(n, mean = 0, sd = 10)

# Combine into a data frame 
data <- data.frame(group = group, score = score)

t.test(data$score, mu = 60) #H0: pop_mean = 50

```

## Two sample test

```{r}
data0 <- filter(data, group == 0) 
data1 <- filter(data, group == 1) 
t.test(data0$score, data1$score) #H0: pop_mean_grp1 = pop_mean_grp2
#OR
t.test(score ~ group, data = data)
```

## NMICS example

```{r}
library(haven)

get_url <- function(x) {
  return(paste0("https://github.com/tempgita/training/raw/refs/heads/master/R%20training%202082-10-03/data/",x))
}

df <- read_spss(get_url('009-hl.sav'))

# HL6 -> Age, HL4 -> Sex 
# Summarize age for males (HL4 == 1) and females (HL4 == 2) 
df_male <- filter(df, HL4 == 1) 
df_female <- filter(df, HL4 == 2) 
mean(df_male$HL6) 
mean(df_female$HL6) 
#average age : male = 28.26, female = 28.83 

t.test(HL6 ~ HL4, data = df) 

summary(lm(HL6 ~ HL4, data = df))
```

# Using Survey Weights

```{r}
library(survey) #for survey data 
library(haven) #for importing Stata dta file 
library(srvyr) # srvyr' for tidyverse-style coding (easier to read) 
 
#function for directly accessing the data from the web 
get_url <- function(x) { 
  return(paste0("https://github.com/tempgita/training/raw/refs/heads/master/R%20training%202082-10-03/data/",x)) 
} 
 
df <- read_stata(get_url('013-poverty.dta')) 

svy_design <- svydesign(ids = ~psu_number,
          strata = ~domain,
          weights = ~hhs_wt,
          data = df)

svy_obj <- as_survey(svy_design)

```

## Population size calculation

```{r}

#manual way
df %>% 
  # group_by(prov) %>% 
  summarise(hh_pop = sum(hhs_wt),
            ind_pop = sum(ind_wt))


svy_obj %>% 
  group_by(prov) %>% 
  summarise(pop_hh = survey_total())
```

## Mean (using survey weight)

```{r}
# The manual way
df %>% 
  summarise(pcep_mean = weighted.mean(pcep, hhs_wt))

#formal way
svy_obj %>% 
  group_by(prov) %>% 
  summarise(pcep_mean = survey_mean(pcep))
```

## Poor proportion

```{r}

svy_obj %>% 
  group_by(prov) %>%
  summarise(poverty_rate = survey_mean(poor,  
                                       vartype = c("se", "ci"), level = 0.95)) # for CI at 5% level of
```

## Poor proportion (individual)

```{r}
ind_svy_design <- svydesign(ids = ~psu_number,
          strata = ~domain,
          weights = ~ind_wt,
          data = df)

ind_svy_obj <- as_survey(ind_svy_design)

ind_svy_obj %>% 
  # group_by(prov) %>%
  summarise(poverty_rate = survey_mean(poor))
```

## Quantiles

```{r}
# --- Using the Survey Object --- 
# Calculate Median PCEP by Province 
svy_obj %>% 
  group_by(prov) %>%  
  summarise( 
    median_pcep = survey_median(pcep), 
    q90 = survey_quantile(pcep, quantiles = 0.9) 
  ) 
```

## Ratio calculation

```{r}
svy_obj %>% 
  group_by(prov) %>%  
  summarise( 
    food_share = survey_ratio(numerator = pcep_food, 
                              denominator = pcep) 
  )
```

## Survey t-test 

```{r}

svyttest(pcep ~ poor, design = svy_obj)
```

## Task: perform t-test between the mean pcep between Madhesh vs Karnali and Madhesh vs. Bagamati. 

```{r}
sub_obj <- svy_obj %>%  
  filter(prov %in% c(6, 7)) %>% 
  mutate(prov_binary =as.factor(prov)) #to make prov binary variable

svyttest(pcep ~ prov_binary, design = sub_obj)
```

# Webscrapping

```{r}
dx <- read.csv("https://data.ny.gov/api/views/d6yy-54nr/rows.csv") 
head(dx)
```

```{r}
# install.packages('rvest')

library(rvest)
library(dplyr) 
 
#loading webpage content 
webpage <- read_html("https://www.sharesansar.com/today-share-price")

tables <- html_table(webpage)
length(tables)

df <- tables[[1]]

df %>% View()
```
